---
title: "Joining Gapminder Data"
author: "Jess Robinson"
output: github_document
---

# Get the original data

```{r, echo = TRUE}
library(tidyverse)
library(gapminder)
library(here)
theme_set(theme_minimal())
gapminder
```

# Import the HIV data

```{r}
#loading HIV data
hivdata <-read_csv(file = here("data", "gapminder_hiv.csv"))
```

# What is the relationship between HIV prevalence and life expectancy?

```{r}
#join data
hivprevlife <- inner_join(hivdata, gapminder)
#plot data
ggplot(data = hivprevlife, mapping = aes(x=hivRate, y=lifeExp)) +
    scale_x_log10() +
    geom_point() +
    geom_smooth() +
    xlab("HIV Rate (log)") +
    ylab("Life Expectancy")
```

For this question, I chose to use an inner join, because I wanted to see only observations where there was data on both HIV prevalence and life expectancy. 
For countries with an logarithmic HIV rate above 0.1, the relationship between HIV prevalence and life expectancy is negative, with higher rates of HIV prevalence associated with lower life expectancies. 

# Which continents have the most countries with missing HIV data?

```{r}
#anti-joining data
missinghiv <- anti_join(gapminder, hivdata)
#plotting missing data
missinghiv %>%
 count(continent) %>%
  filter(!is.na(continent)) %>%
  ggplot(mapping = aes(x = reorder(continent, -n), y=n, fill=continent)) +
    geom_col() +
    labs(title = "Missing HIV data by Continent",
    x= "Continents",
    y = "Number of Missing Observations",
    fill = "Continents")
```

For this question, I chose to use an anti-join, because I wanted to see observations that had missing HIV data. 
Africa, followed by Asia and then Europe, have the greatest number of missing HIV data observations. 





## Session info

```{r, echo = TRUE}
devtools::session_info()
```


